---
output:
  html_document:
    smart: no
  variant: default
---

Toronto Water
===========================================================================

Susan Li
April 2, 2017

I have been exploring [Toronto's Open Data Catalogue](http://www1.toronto.ca/wps/portal/contentonly?vgnextoid=1a66e03bb8d1e310VgnVCM10000071d60f89RCRD), but I have to admit that it is not always easy to find interesting dataset. Today, this [water consumption by ward 2000 to 2015 dataset](http://www1.toronto.ca/wps/portal/contentonly?vgnextoid=289bd103cd8b1310VgnVCM1000003dd60f89RCRD) got my attention. To put things in perspective, I will only look into the data for the past five years.

```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE)
```

As usual, download dateset, check missing values and do some wrangling. There are five columns that I do not need, I decided to remove them.

```{r}
library(xlsx)
library(ggplot2)
library(ggthemes)
library(gridExtra)
library(reshape2)
water2011 <- read.xlsx('water_consumption_2011.xls', sheetIndex = 1)
water2012 <- read.xlsx('water_consumption_2012.xls', sheetIndex = 1)
water2013 <- read.xlsx('water_consumption_2013.xls', sheetIndex = 1)
water2014 <- read.xlsx('water_consumption_2014.xls', sheetIndex = 1)
water2015 <- read.xlsx('water_consumption_2015.xls', sheetIndex = 1)
water <- rbind(water2011, water2012, water2013, water2014, water2015)
water$year <- as.numeric(as.character(water$year))
water <- water[complete.cases(water[,-1]),]
```

```{r}
water$residential.accounts <- NULL
water$total.consumption <- NULL
water$average.consumption <- NULL
water$commercial.accounts <- NULL
water$total.count <- NULL
```

```{r}
options("scipen"=100, "digits"=4)
p1 <- ggplot(aes(x = city.ward, y = annual.residential.usage), data = water) +
  geom_line(size = 2) +
  scale_x_continuous(breaks = seq(1, 45, 1)) +
  ylab('Water Usage(in cubic meters)') +
  ggtitle('Total Residential Usage by Ward 2011 - 2015') +
  theme_economist()
p2 <- ggplot(aes(x = city.ward, y = annual.commercial.usage), data=water) +
  geom_line(size = 2) +
  scale_x_continuous(breaks = seq(1, 45, 1)) +
  ylab('Water Usage(in cubic meters)') +
  ggtitle('Total Commercial Usage by Ward 2011 - 2015') +
  theme_economist()
grid.arrange(p1, p2, ncol=1)
```

It is clear that ward 42, - Scarborough-Rouge River had the most residential usage and Ward 27, - Toronto Centre Rosedale had the most commercial usage in total from 2011 to 2015. 

```{r}
p3 <- ggplot(data=water,
       aes(x=city.ward, y=average.residential.usage)) +
       geom_line(size = 2) +
  scale_x_continuous(breaks = seq(1, 45, 1)) +
  ylab('Water Usage(in cubic meters)') +
  ggtitle('Average Residential Usage by Ward 2011 - 2015') +
  theme_economist()
p4 <- ggplot(data=water,
       aes(x=city.ward, y=average.commercial.usage)) +
       geom_line(size = 2) +
  scale_x_continuous(breaks = seq(1, 45, 1)) +
  ylab('Water Usage(in cubic meters)') +
  ggtitle('Average Commercial Usage by Ward 2011 - 2015') +
  theme_economist()
grid.arrange(p3, p4, ncol=1)
```

Again, ward 27, - Toronto Centre Rosedale had the most residential usage on average from 2011 to 2015, and ward 11, - York South West consumed the most water commercially on average from 2011 to 2015. 

```{r}
water1 <- water
water1$average.residential.usage <- NULL
water1$average.commercial.usage <- NULL
water_long <- melt(water1, id = c('city.ward', 'year'))
ggplot(data=water_long,
       aes(x=city.ward, y=value, color = variable)) +
       geom_line() +
  ylab('Total Water Usage(in cubic meters)') +
  ggtitle('Total Water Usage') + facet_wrap(~year) + theme_economist()
```

It has been consistent that ward 27 consumed the most water commercially every year from 2011 to 2015. 

```{r}
ggplot(data=water, 
       aes(x=city.ward, y=average.commercial.usage)) +
       geom_line() +
  ylab('Water Usage(in cubic meters)') +
  ggtitle('Average Commercial Usage') + facet_wrap(~year) + theme_economist()
```

```{r}
ggplot(data=water,
       aes(x=city.ward, y=average.residential.usage)) +
       geom_line() +
  ylab('Water Usage(in cubic meters)') +
  ggtitle('Average Residential Usage') + facet_wrap(~year) + theme_economist()
```

The average residential usage has been decreasing over the years, we now use less water then we did five years ago. 

```{r}
water2011 <- water[water$year==2011 & water$average.residential.usage, ] 
water2015 <- water[water$year==2015 & water$average.residential.usage, ]
summary(water2011$average.residential.usage)
summary(water2015$average.residential.usage)
```

Now what? It's map time! I downloaded the ESRI Shapefile from [city of Toronto Open Data portal](http://www1.toronto.ca/wps/portal/contentonly?vgnextoid=b1533f0aacaaa210VgnVCM1000006cd60f89RCRD&vgnextchannel=1a66e03bb8d1e310VgnVCM10000071d60f89RCRD). This data set includes the boundaries for the City of Toronto's 44 municipal wards.

```{r}
# report shapefile details
library(rgdal)
library(sp)
ogrInfo("C:/Users/Susan/Documents/gcc/Projects/Open Data/Files/Data Upload - May 2010/May2010_WGS84", "icitw_wgs84")
```

It took me sometime to figure it out how to read it. 

```{r}
# Read in shapefile 
toronto.rg <- readOGR("C:/Users/Susan/Documents/gcc/Projects/Open Data/Files/Data Upload - May 2010/May2010_WGS84", "icitw_wgs84")
print(proj4string(toronto.rg))
```

```{r}
# Generate a simple map to show all three layers
plot(toronto.rg, axes=TRUE, border="gray")
```

Here we have a simple map of Toronto to show all three layers.

Use 'fortify' function to turn the map into a data frame so that it can easily be plotted with 'ggplot2', which produce this data frame:

```{r}
library(rgeos)
toronto.rg.df <- fortify(toronto.rg, region = "SCODE_NAME")
toronto.rg.df$id <- as.integer(toronto.rg.df$id)
ggplot(data=toronto.rg.df, aes(x=long, y=lat, group=group)) + 
  geom_polygon(fill = 'grey80', color = 'black', size = 1) + theme()
head(toronto.rg.df)
```

Merge our water dataframe with the map data. This time, I will be looking at 2015 usage only.

```{r}
water_2015_df <- merge(toronto.rg.df, water2015, by.x = 'id', by.y = 'city.ward')
water_2015_df <- water_2015_df[order(water_2015_df$order), ]
# get the centroids and then convert them to a SpatialPointsDataFrame for ward labelling
centroids  <- as.data.frame(gCentroid(toronto.rg, byid = T))
centroids[['id']] <- toronto.rg@data$SCODE_NAME
```

Plot 2015 total commercial usage!

```{r}
total_commercial_2015 <- water_2015_df[c(1:8,11)]

ggplot(data = total_commercial_2015, aes(x = long, y = lat, group = group)) + 
  geom_polygon(aes(fill = annual.commercial.usage), color = 'red', alpha = 0.8) +
  scale_fill_gradient(name='Water Usage in Cubic Meters',low = '#56B1F7', high = "#132B43") +
  labs(title = "Toronto Total Commercial Water Usage by Ward 2015") +
  geom_text(aes(x=x,y=y, group=NULL, label=id), data = centroids, size=2)
```

2015 Total Residential Usage

```{r}
total_residential_2015 <- water_2015_df[c(1:9)]

ggplot(data = total_residential_2015, aes(x = long, y = lat, group = group)) + 
  geom_polygon(aes(fill = annual.residential.usage), color = 'red', alpha = 0.8) +
  scale_fill_gradient(name='Water Usage in Cubic Meters',low = '#56B1F7', high = "#132B43") +
  labs(title = "Toronto Total Residential Water Usage by Ward 2015") +
  geom_text(aes(x=x,y=y, group=NULL, label=id), data = centroids, size=2)
```

Plot 2015 Average Commercial Water Usage.

```{r}
average_commercial_2015 <- water_2015_df[c(1:8, 12)]
ggplot(data = average_commercial_2015, aes(x = long, y = lat, group = group)) + 
  geom_polygon(aes(fill = average.commercial.usage), color = 'red', alpha = 0.8) +
  scale_fill_gradient(name='Water Usage in Cubic Meters', low = '#56B1F7', high = "#132B43") +
  labs(title = "Toronto Average Commercial Water Usage by Ward 2015") +
  geom_text(aes(x=x,y=y, group=NULL, label=id), data = centroids, size=2)
```

2015 Average Residential Usage

```{r}
average_residential_2015 <- water_2015_df[c(1:8, 10)]
ggplot(data = average_residential_2015, aes(x = long, y = lat, group = group)) +
  geom_polygon(aes(fill = average.residential.usage), color = 'red', alpha = 0.8) +
  scale_fill_gradient(name='Water Usage in Cubic Meters', low = '#56B1F7', high = "#132B43") +
  labs(title = "Toronto Average Residential Water Usage by Ward 2015") +
  geom_text(aes(x=x,y=y, group=NULL, label=id), data = centroids, size=2)
```

## The end

Commercial water users are the top consumers of the most water in Toronto, and they are spread out throughout the city in many wards. And heavy residential water users are only concentrated in a few wards of the city.

Water use is a major policy, environmental and social issue. Understanding water usage is important everywhere. The cost of water went up in January of 2014 by nine per cent. We are lucky that [Toronto is built on a lake that contains about 1640 cubic kilometers of water](http://globalnews.ca/news/1672371/how-much-water-does-toronto-use-each-day/). But [New study calls average water use by Canadians 'alarming'](http://www.nationalpost.com/rss/study+calls+average+water+Canadians+alarming/1402591/story.html).

I have really enjoyed working on this project. I hope the city of Toronto will release the data by month, and type of user such as residence, park, business, school, hotel, restaurant etc, so that I will be able to perform a more in depth analysis. 
